---
title: "kable Examples"
output: html_notebook
---

# Example usage of kables

```{r}
library(knitr)
library(kableExtra)
```

Consider, first, the `Auto` dataframe:

```{r}
Auto
```

The basic `kable` formatted version is:

```{r}
kable(Auto)
```

Notice that this looks better, but it includes a column on the left with row numbers, which we may not want. We can also add a table title caption:

```{r}
kable(Auto, row.names = FALSE, caption = "Auto data formatted as a kable")
```

Tweaks to the appearance can be generated with the `kableExtra` package. the following uses the `kable_styling` style, makes better horizontal use of space, and alternates striping of the rows:

```{r}
kb1 <- kable(Auto, row.names = FALSE, caption = "Auto data formatted as a kable")
kable_styling(kb1, full_width =FALSE, bootstrap_options = "striped")
```

As another example:

```{r}
kb1 <- kable(Auto, row.names = FALSE, caption = "Auto data formatted as a kable")
kable_classic(kb1, full_width =FALSE, bootstrap_options = "striped", font_size = 10)
```

Additional details on options and styles can be found at:

-   <https://bookdown.org/yihui/rmarkdown-cookbook/kable.html>

-   <https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html>

amongst other websites. You are encouraged to use such output
