---
title: "3/15/2023 Lecture"
output: html_notebook
---

```{r}
library(tidyverse)
library(mvtnorm) # for multivariate normals
library(ggfortify) # for PCA visualization
```

# Unsupervised Learning Problems

How do we detect:

-   Directions along which variations are greatest

-   Clusters

```{r}
set.seed(100)
mu <- c(3,-5)
sigma <- matrix(c(2, 1, 1, 2), nrow=2)
xy.samples <- as_tibble(rmvnorm(10^3, mean = mu, sigma = sigma))
colnames(xy.samples) <- c("x","y")
xy.plt <- ggplot(data=xy.samples, mapping = aes(x=x,y=y)) +geom_point()
print(xy.plt)
set.seed(100)
mu1 <- c(2,-4)
sigma1 <- matrix(c(2, 1, 1, 2), nrow=2)
mu2 <- c(-1,2)
sigma2 <- matrix(c(3, -0.1, -0.1, 3), nrow=2)
samples1 <- rmvnorm(10^3, mean = mu1, sigma = sigma1)
samples2 <- rmvnorm(10^3, mean = mu2, sigma = sigma2)
clusters.samples <- as_tibble(rbind(samples1, samples2))
colnames(clusters.samples) <- c("x","y")
clusters.plt <- ggplot(data=clusters.samples, mapping = aes(x=x,y=y)) + 
  geom_point()
print(clusters.plt)
```

# PCA

## Run PCA on the skewed normal

```{r}
pr.out <- prcomp(xy.samples)
pr.out
```

Squares of the standard devs are the eigenvalues:

```{r}
pr.out$sdev^2
```

Proportion of variance explained (PVE):

```{r}
pr.var <- pr.out$sdev^2
pve <- pr.var/sum(pr.var)
pve
```

Suppose we scale our data (in general, we should):

```{r}
pr.scaled.out <- prcomp(xy.samples, scale=TRUE)
pr.scaled.out
```

```{r}
pr.scaled.var <- pr.scaled.out$sdev^2
pve.scaled <- pr.scaled.var/sum(pr.scaled.var)
pve.scaled
```

```{r}
names(pr.scaled.out)
```

```{r}
pr.scaled.out$center
```

## Biplot visualization

See how data is distributed along the PCA1-PCA2 axes:

```{r}
pca.scaled.plt <- autoplot(pr.scaled.out) + 
  ggtitle("PCA Scaled/No Loadings")
print(pca.scaled.plt)

pca.scaled.loadings.plt <- autoplot(pr.scaled.out,loadings=TRUE,
                                    loadings.label=TRUE) + 
  ggtitle("PCA Scaled/With Loadings")
print(pca.scaled.loadings.plt)
```

## Iris Data

This runs PCA on <https://archive.ics.uci.edu/ml/datasets/iris>, the iris data set.

```{r}
iris.df <- read_csv("../data/iris.data",col_names=FALSE, show_col_types = FALSE)
colnames(iris.df) <- c("sepal.length", "sepal.width", 
                       "petal.length", "petal.width", "class")
iris.df$class <- as_factor(iris.df$class)

# compute principle components
pca.iris <- prcomp(select_if(iris.df,is.numeric),scale=TRUE)
pr.var <- pca.iris$sdev^2
pve <- pr.var/sum(pr.var)

pve
```

```{r}
iris.plt <- autoplot(pca.iris,loadings=TRUE,loadings.label=TRUE) +  ggtitle("PCA of Iris Data")
print(iris.plt)
```
