---
title: "Lecture 1/11/2023"
output: html_notebook
---

# Generating Random Variables

Example of sampling the $U(0,1)$, uniform, distribution

```{r}
runif(1)
runif(10)
x <- runif(5)
x
```

For $U(a,b)$,

```{r}
a <- 1
b <- 5
x <- runif(5,min = a, max= b)
x
y <- runif(5, min=1, max= 5)
y
```

Reproducibility of random numbers can be achieved by settting the **seed**:

```{r}
set.seed(123)
a <- 1
b <- 5
x <- runif(5,min = a, max= b)
x
set.seed(123) # reset seed before calling again
y <- runif(5, min=1, max= 5)
y
```

In any code involving random numbers, set a seed explicitly at the beginning.

Gaussian random variable with mean $m$ and variance \$\\sigma\^2\$, $N(m, \sigma^2)$:

```{r}
set.seed(123)
rnorm(1) #defaults to m = 0, sigma = 1
rnorm(10)
rnorm(10, mean = 5, sd = .1)
```

**NOTE** R parameterizes `rnorm` by mean and standard deviation, not variance.

To call an R script, use the `source` command:

```{r}
source("~/code/Math_318_2023/scripts/script1.R")
getwd()
```

```{r}
getwd()
```

```{r}
setwd("~/code/")
getwd()
```

Elementary statistics:

```{r}
set.seed(100)
xdata <- rnorm(100)
mean(xdata)
var(xdata)
summary(xdata)
quantile(xdata)
quantile(xdata, .3) # get the 30% quantile
```
