---
title: "2/10/23 Lecture"
output: html_notebook
---

# Categorical Response Variables

```{r}
library(tidyverse)
library(ISLR2)
```

Look at the `Default` dataset, part of `ISLR2`.

```{r}
Default <- as_tibble(Default)
Default
```

This is credit card/financial information. Want to predict if a customer will default, based on if they are a student (or not), their income, and their credit card balance.

Visualize defaults

```{r}
box_plt <- ggplot(Default, aes(default, balance)) + geom_boxplot()
print(box_plt)
```

Does income play a role?

```{r}
box_plt <- ggplot(Default, aes(default, income)) + geom_boxplot()
print(box_plt)
```

Try to use linear regression: use built in `ggplot2` for simplicity

```{r}
lm_plt <- ggplot(Default, aes(x=balance, y = as.numeric(default)-1)) + geom_point() +
  geom_smooth(method=lm, formula = y~x) + xlab("Balance") + ylab("Default Probability") +
  ggtitle("Default Data")
print(lm_plt)
```

```{r}
as.numeric(Default$default)-1 # check what this is doing
```

Train a logistic classifier using `glm`

```{r}
default_classifier <- glm(default~balance,data=Default, family = binomial)
```

Returns $\hat{\beta}_0$ and $\hat{\beta}_1$ to get the default classifier

$$
\mathbb{P}(Default\mid balance) = \phi(\hat{\beta}_0 + \hat{\beta}_1\times balance)
$$

```{r}
default_classifier
```

```{r}
summary(default_classifier)
```
