---
title: "Lecture 1/13/2023"
output: html_notebook
---

# Introducing tibbles and control structures

```{r}
library(tidyverse)
cast <- tibble("Names" = c("Jerry", "George", "Elaine", "Cosmo"), 
               "Ages"=c(38, 38, 37,999))
cast
```

Change the 999 to `NA` , "not available"

```{r}
for (j in 1:nrow(cast)){
  if(cast$Ages[j]>100){
    cast$Ages[j] <- NA
  }
}
print(cast)
```

# Loading in data

```{r}
library(readxl) # part of tidyverse
```

```{r}
getwd()
#grades <- read_excel("../data/Grades.xlsx")
grades <- read_excel("/Users/grs53/code/Math_318_2023/data/Grades.xlsx")
print(grades)
```

Tibble is a type of **data frame**:

```{r}
print(grades[1,])
print(grades[2:5,])
```

`$names` are the names of the columns in the data frame

```{r}
attributes(grades) # find out details of the data frame
print(grades$Grade)
```

Manipulations:

```{r}
summary(grades)
```

By default, R is treating the student ID as a numeric value; it is really categorical

```{r}
grades$Student <- factor(grades$Student) # convert to categorical type
print(grades)
summary(grades)
```

CSV, comma separated value, data is common

```{r}
grades.csv <- read.csv("../data/Grades.csv")
print(grades.csv) # default data frame type
grades.df <- as_tibble(grades.csv) # casts to tibble type
print(grades.df)
```

# Getting data out of R

```{r}
x <- seq(0,1,length.out=11)
print(x)
write.csv(x,"xdata.csv")
```

```{r}
set.seed(100)
x <- runif(11)
x.summary <- summary(x)
write(x.summary, file="summary.txt")
```

# More manipulations of data sets

```{r}
library(ISLR2)
print(Auto)
```

```{r}
Auto <- as_tibble(Auto)
Auto
```

```{r}
Auto$cylinders <- factor(Auto$cylinders)
summary(Auto$cylinders)
```

```{r}
Auto$origin <- factor(Auto$origin)
summary(Auto$origin)
```

```{r}
Auto
```

# Intro to graphics

```{r}
x <-  seq(0,1,length.out=51)
y <- x * (1-x) # (x,y) data set
plot(x,y)
```

Dump to pdf: This plots the figure, first to the screen, and then opens a file, and saves the output to disk.

```{r}
plot(x,y, main ="Example Plot", xlab="x",  ylab="y", type = "l", col="blue", lwd=2)
pdf(file="plot_example1.pdf") # open file for saving graphics
plot(x,y, main ="Example Plot", xlab="x",  ylab="y", type = "l", col="blue", lwd=2)
dev.off() # close graphics file
```
