---
title: "2/3/2023 Lecture"
output: html_notebook
---

# Multilinear Regression

```{r}
library(tidyverse)
```

## Simple Example

Create some data,

$$
Y = 5 + 2 X_1 -3 X_2 + \epsilon
$$

```{r}
n = 15 # number of observations
beta0 = 5
beta1 = 2
beta2 = -3 # true coefficients

x1 <-  runif(n, min=-2, max=2) # create random measurement points
x2 <-  runif(n, min=-2, max=2)
# data
y = beta0+beta1*x1 + beta2*x2 + rnorm(n) # add some noise
y
```

Fit the model

```{r}
df <- tibble(x1=x1, x2=x2, y=y) # package data into a tibble

multi <- lm(y~ x1 + x2,df) # regress
```

```{r}
print(multi)
```

Report confidence intervals (95%) for the coefficients

```{r}
confint(multi)

```

## Multilinear regression in the Auto dataset

```{r}
library(ISLR2)
```

Look at pair correlations for insight in what might relate to what:

```{r}
cor(select_if(Auto, is.numeric))
```

```{r}
cor.df <- cor(select_if(Auto, is.numeric))
```

```{r}
cor.df <- as_tibble(cor.df)
```

```{r}
cor.df
```

```{r}
abs(cor.df["mpg"])>0.8
```

```{r}
names(cor.df)[abs(cor.df["mpg"])>0.8]
```

Try a linear model using displacement, horsepower and weight, to predict MPG. But lets do a different fit for each number of cylinders and report results:

```{r}
for (cyl in sort(unique(Auto$cylinders))){
  cat("Case of cylinders = ", cyl)
  # fit horsepower, weight, displacement, and segment data cylindres, from Auto dataframe
  auto_lm <- lm(mpg~horsepower+weight+displacement, I(cylinders==cyl), data=Auto)
  print(summary(auto_lm))
}
```

Look at a particular case

```{r}
lm(mpg~horsepower+weight+displacement, data=filter(Auto, cylinders==4))
```
