---
title: "1/25/2023 Lecture"
output: html_notebook
---

# Confidence Intervals

```{r}
set.seed(100)
x <- rnorm(10) # 10 N(0,1) random numbers
mean(x)
```

How confident are we in that representing the population mean?

```{r}
set.seed(100)
x <- rnorm(10^6) # 10^6 N(0,1) random numbers
mean(x)
```

Computing the 95% CI in R:

$$
\left( \bar{x}_n - z_{1-\alpha/2} \frac{\hat{\sigma}_n}{\sqrt{n}}, \bar{x}_n + z_{1-\alpha/2} \frac{\hat{\sigma}_n}{\sqrt{n}} \right)
$$

on $n$ samples. $\bar{x}_n$ is the sample mean. $\hat{\sigma}_n$ is the sample standard dev., and $z_{1-\alpha/2}$ is the \$1-\\alpha/2\$% quantile of $N(0,1)$.

For a given significance level, $\alpha$:

-   Find the sample mean and standard dev

-   Find $z_{1-\alpha/2}$

```{r}
set.seed(100)
x <- rnorm(10^3) # 10^6 N(0,1) random numbers
alpha = 0.05 # corresponds to 95% CI
z = qnorm(1- alpha/2) # this is the z_{1-\alpha/2} score
z

```

```{r}
n = length(x)
ci = c(mean(x) - z * sd(x)/sqrt(n),mean(x) + z * sd(x)/sqrt(n))
ci
```

Built in tool for the CI:

```{r}
t.test(x)
```

```{r}
t.test(x, conf.level=.9)
```

**NOTE**: This is performing a Student's t-test, not the z-test, so there is some deviation from 1.96 for 95% CIs when you have a small sample size.

# More on Data Frames

See Chapters 9 and 10 of <https://r4ds.had.co.nz> for additional details on this section.

```{r}
library(tidyverse)
library(ISLR2) # for the Auto data frame
```

```{r}
Auto
```

Use the `gather` command:

```{r}
Auto.long <- gather(Auto, key="variable", value, -year)
```

```{r}
Auto.long
```

Set it to keep some things numerical:

```{r}

Auto.long.numeric <- gather(Auto, key=variable, value, -cylinders, -year, -origin)
```

```{r}
Auto.long.numeric
```

Mapping data back into columns,

```{r}
Auto.short <- spread(Auto.long, key=variable, value = value)
```

Is not possible

Example of why we would bother:

```{r}
stocks <- tibble(time = as.Date('2009-01-01') + 0:9, 
                 X = rnorm(10, 0,1), 
                 Y = rnorm(10, 0 ,2),
                 Z = rnorm(10,0,4))
stocks
```

At each date, we have the stock "price" for X, Y, and Z. Each row contains **three** observations (X,Y,Z).

```{r}
stocks.long <- gather(stocks, key=symbol, value = price, -time)
stocks.long
```

Now each row is a single measurement. Since time/symbol pairs are unique, this can be undone:

```{r}
stocks.short <- spread(stocks.long, key=symbol, value = price)
stocks.short
```

## Filtering Data

Look at cars only from 1970:

```{r}
Auto.70 <- filter(Auto, year == 70)
summary(Auto.70)
```

Filter by cylinders and year:

```{r}
Auto.subset <- filter(Auto, year> 75, cylinders ==4)
summary(Auto.subset)
```

Filter by columns:

```{r}
Auto.subset <- select(Auto, mpg, cylinders, displacement)
Auto.subset
```

Want statistics for certain subsets of the data. Get the 25- and 75-percentile of MPG, as a function of cylinders:

```{r}
by.cylinders <- group_by(Auto, cylinders) # separates data by cylinders factor
summarize(by.cylinders, "25% Quantile"= quantile(mpg)[2],"75% Quantile"=quantile(mpg)[4])
```

Recall:

```{r}
quantile(Auto$mpg)
```

```{r}
by.cylinders <- group_by(Auto, cylinders) # separates data by cylinders factor
summarize(by.cylinders, "Mean MPG"= mean(mpg))
```

```{r}
by.cylinders
```

## Adding Rows and Columnbs

```{r}
df <- tibble(x=1:3, y = 3:1)
df
```

Add a row

```{r}
add_row(df, x= 5, y = -4)
```

```{r}
add_column(df, sample = "one")
```

## Combining two data frames

```{r}
df1 <- tibble(x=1:3, y = 3:1)
df2 <- tibble(x=runif(3), y = rnorm(3))
union_all(df1, df2)
```

## Cleaning data

Remove `NA` from tables

```{r}
x= seq(1:10)
y <- x^2
y[4]= NA
xy.df <- tibble(x=x,y=y)
```

```{r}
xy.df
```

Clean the observation with missing entry:

```{r}
xy.clean.df <- na.omit(xy.df)
xy.clean.df
```

```{r}
NaN
```
