---
title: "Lecture 1/23/2023"
output:
  html_document:
    df_print: paged
---

# Correlations

```{r}
library(tidyverse)
```

## Toy Example

Create data

$$
y_i = x_i^3 + \epsilon_i
$$

Think of $\epsilon$ as "noise" in the measurements, and for simplicity, assume $\epsilon_i \sim N(0, \sigma^2)$ for some particular variance, $\sigma$ .

```{r}
set.seed(100) # set a seed for reproducibility
x <- seq(-5, 5, length=101) # 101 uniformly spaced points between -5 and 5
xi <-  rnorm(length(x)) # create a vector of Gaussians with same size as x, N(0,1)

# look at different magnitude noises
sigma2_vals = c(0, 10, 50,100, 500) 

# store in a data structure, data frame, with the columns we want
df <- tibble(x = numeric(), y=numeric(), noise=factor())
df

```

```{r}
#populate the data frame
for(j in 1:length(sigma2_vals)){
  sigma2 = sigma2_vals[j] # get the jth coordinate
  y <- x^3 + (sigma2) * xi # add noise with variance sigma2 to x^3
  df <- add_row(df, x = x, y = y, noise = as.factor(sigma2))
}
df
```

Compute the data correlations between $x$ and $y$ at the different noise levels

```{r}
# vector of correlations at the three different noise levels
correlations=c() # empty to start
for(j in 1:length(sigma2_vals)){
  sigma2 = sigma2_vals[j] # look at just the j-th noise case
  correlations <- append(correlations, 
                         cor(filter(df, noise == sigma2)$x,
                             filter(df, noise == sigma2)$y))
}
correlations
```

Look at this visually

```{r}
plt <- ggplot(df) + geom_point(aes(x,y, color=noise))
plt <- plt + scale_color_discrete(name="Noise Intensity")
print(plt)
```

\
Example of uncorrelated data: what to expect graphically:

```{r}
z <-  runif(length(x))
df2 <- tibble(x = x, z =z)
plt2 <- ggplot(df2)+ geom_point(aes(x, z))+ ggtitle("Example of Uncorrelated Data")
print(plt2)
```

$Z$ is independent of $X$.

## Correlations in a real data set

Look at the `Auto` data set.

```{r}
library(ISLR2)
Auto
```

Set categorical variables

```{r}
Auto$origin <- as_factor(Auto$origin)
Auto$cylinders <- as_factor(Auto$cylinders)
Auto$year <- as_factor(Auto$year)
```

Quickly look at all the pair correlations:

```{r}
# filter out non-numeric data types
cor(select_if(Auto, is.numeric))
```

Correlation pairs are symmetric

$$
\text{Cor}(X,Y) = \text{Cor}(Y,X)
$$\
Diagonal is always 1 since

$$
\text{Cor}(X, X) = 1
$$

Quick observations:

-   Larger engine correlates with more horsepower, and heavier vehicle

-   Heavier vehicle negatively correlates with MPG

Graphical version of this

```{r}
library(GGally)
plt <- ggpairs(select_if(Auto, is.numeric))+ggtitle("Pair Correlations for Auto")
print(plt)
```
